Preventing runtime errors of Redis at compile time

### 2015-11-18 paper.tex
draft

Nonetheless, it's still useful to have a term-level value that links with a
Symbol, when we want to retreive type-level informations at runtime (but not the
 other way around!).

@@

With all of these ingrediants ready, let's build some dictionaries!

@@

\emph{Indexed monads} (or \emph{monadish},
\emph{parameterised monad})\cite{indexedmonad}
can be used\cite{typefun}\cite{staticresources} to model such preconditions and
postconditions in types. An indexed monad is a type constructor that takes three
arguments: an intitial state, a final state, and the type of a value that an
action computes, which can be read like a Hoare triple\cite{kleisli}.

@@

We define a new datatype \mintinline{haskell}{Edis}, which is basically just
\mintinline{haskell}{Redis} indexed with more informations in types. And we make
 it an instance of \mintinline{haskell}{IMonad}.

@@

To execute a \mintinline{haskell}{Edis} program, simply apply it to
 \mintinline{haskell}{unEdis} to get an ordinary
 \mintinline{haskell}{Hedis} program, with type informations erased.

@@

Dicionary \mintinline{haskell}{xs} in the type remain unaffected after the
action, because \mintinline{haskell}{ping} does not affect any key-value
bindings. To encode other commands that alters key-value bindings, we need
 type-level functions to annotate those effects on the dictionary.

@@

Type families have a wide variaty of applications. They can appear inside type
classes\cite{tfclass}\cite{tfsynonym}, or at toplevel. Toplevel type families
can be used to compute over types, they come in two forms: open\cite{tfopen} and
closed \cite{tfclosed}.

@@

The first instance of \mintinline{haskell}{And} could be subsumed
under the more general instance \mintinline{haskell}{And a b}.
But the closedness allows these instances to be resoluted in order, just like
how cases are resoluted in term-level functions. Also notice that how much
\mintinline{haskell}{And} resembles to it's term-level sibling.

@@

Another benifit of closed type families is that type-level equality can be
expressed by unifying type variables with the same name.
\mintinline{haskell}{Get} takes two type arguments, a dictionary and a symbol.
If the key we are looking for unifies with the symbol of an entry, then
 \mintinline{haskell}{Get} returns the corresponding type, else it keeps
 searching down the rest of the dictionary.

@@

\mintinline{haskell}{KnownSymbol} is a class that gives the string associated
with a concrete type-level symbol, which can be retreived with
\mintinline{haskell}{symbolVal}.\footnotemark

@@

Since Haskell has a \emph{phase distinction}\cite{phasedistinction}, types are
erased before runtime. It's impossible to obtain informations directly from
types, we can only do this indirectly, with \emph{singleton types}
\cite{singletons}.

@@

In the type of \mintinline{haskell}{del}, the type varabile
\mintinline{haskell}{s} would be a \mintinline{haskell}{Symbol} that is decided
 by the argument of type \mintinline{haskell}{Proxy s}.
To use \mintinline{haskell}{del}, we would have to apply it with a clumsy
 term-level proxy like this:

@@

After \mintinline{text}{SET}, the key will be associated with
\mintinline{haskell}{Strings} in the dictionary, indicating that it's a string.
%
% So far, the datatypes we are dealing with, are all of those built-in Redis data
% structures, such as \emph{strings}, \emph{lists (of strings)},
% \emph{sets (of strings)}, etc.

@@

\footnotetext{The methods of \mintinline{haskell}{Serialize} will have default
generic implementations for all datatypes with some language extenstions
enabled, no sweat!}

\subsection{Extending container types}
We rename container types and extend it with an extra type arguement,
 to indicate what kind of encoded value it's holding.

@@

@@

Consider the following example: \mintinline{text}{LLEN} retrurns the length of
the list associated with a key, else raises a type error.

@@

We could achieve this simply by translating the semantics we want to the
domain of Booleans, with type-level boolean functions such as
\mintinline{haskell}{(&&)},
\mintinline{haskell}{(||)}, \mintinline{haskell}{Not},
\mintinline{haskell}{(==)}, etc.\footnotemark To avoid

@@

To avoid adressing the type of value (as it may not exist at all), we defined an
 auxiliary predicate \mintinline{haskell}{IsList :: Maybe * -> Bool} to replace
 the former part.

@@

With \emph{constraint kind}, a recent addition to GHC. Type constraints now has
its own kind: \mintinline{haskell}{Constraint}. That means type constraints
are not restricted to the left side of a \mintinline{haskell}{=>} anymore,
they could appear in anywhere that accepts something of kind
\mintinline{haskell}{Constraint}, and any type that has kind
\mintinline{haskell}{Constraint} can also be used as a type constraint.\footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}
We could give the lengthy constraint above a type synonym.

@@

\section{Discussions}

\section{Conclusions and Related Work}

>>>>

Nonetheless, it's still useful to have a term-level value that links with a
Symbol, when we want to retrieve type-level information at runtime (but not the
 other way around!).

@@

With all of these ingredients ready, let's build some dictionaries!

@@

\emph{Indexed monads} (or \emph{monadish},
\emph{parameterised monad})\cite{indexedmonad}
can be used\cite{typefun}\cite{staticresources} to model such preconditions and
postconditions in types. An indexed monad is a type constructor that takes three
arguments: an initial state, a final state, and the type of a value that an
action computes, which can be read like a Hoare triple\cite{kleisli}.

@@

We define a new datatype \mintinline{haskell}{Edis}, which is basically just
\mintinline{haskell}{Redis} indexed with more information in types. And we make
 it an instance of \mintinline{haskell}{IMonad}.

@@

To execute a \mintinline{haskell}{Edis} program, simply apply it to
 \mintinline{haskell}{unEdis} to get an ordinary
 \mintinline{haskell}{Hedis} program, with type information erased.

@@

Dictionary \mintinline{haskell}{xs} in the type remain unaffected after the
action, because \mintinline{haskell}{ping} does not affect any key-value
bindings. To encode other commands that alters key-value bindings, we need
 type-level functions to annotate those effects on the dictionary.

@@

Type families have a wide variety of applications. They can appear inside type
classes\cite{tfclass}\cite{tfsynonym}, or at toplevel. Toplevel type families
can be used to compute over types, they come in two forms: open\cite{tfopen} and
closed \cite{tfclosed}.

@@

The first instance of \mintinline{haskell}{And} could be subsumed
under the more general instance \mintinline{haskell}{And a b}.
But the closedness allows these instances to be resolved in order, just like
how cases are resolved in term-level functions. Also notice that how much
\mintinline{haskell}{And} resembles to it's term-level sibling.

@@

Another benefit of closed type families is that type-level equality can be
expressed by unifying type variables with the same name.
\mintinline{haskell}{Get} takes two type arguments, a dictionary and a symbol.
If the key we are looking for unifies with the symbol of an entry, then
 \mintinline{haskell}{Get} returns the corresponding type, else it keeps
 searching down the rest of the dictionary.

@@

\mintinline{haskell}{KnownSymbol} is a class that gives the string associated
with a concrete type-level symbol, which can be retrieved with
\mintinline{haskell}{symbolVal}.\footnotemark

@@

Since Haskell has a \emph{phase distinction}\cite{phasedistinction}, types are
erased before runtime. It's impossible to obtain information directly from
types, we can only do this indirectly, with \emph{singleton types}
\cite{singletons}.

@@

In the type of \mintinline{haskell}{del}, the type variable
\mintinline{haskell}{s} would be a \mintinline{haskell}{Symbol} that is decided
 by the argument of type \mintinline{haskell}{Proxy s}.
To use \mintinline{haskell}{del}, we would have to apply it with a clumsy
 term-level proxy like this:

@@

After \mintinline{text}{SET}, the key will be associated with
\mintinline{haskell}{Strings} in the dictionary, indicating that it's a string.

@@

\footnotetext{The methods of \mintinline{haskell}{Serialize} will have default
generic implementations for all datatypes with some language extensions
enabled, no sweat!}

\subsection{Extending container types}
We rename container types and extend it with an extra type argument,
 to indicate what kind of encoded value it's holding.

@@

\subsection{Handling \mintinline{text}{INCR}}

Commands such as \mintinline{text}{INCR}, \mintinline{text}{INCRBYFLOAT}, are
 not only container-specific, there are also some requirements on what types of
 value they could operate on.

We handle this by mapping Redis's strings of integers and floats to Haskell's
\mintinline{text}{Integer} and \mintinline{text}{Double}.

\begin{minted}[xleftmargin=1em,fontsize=\footnotesize]{haskell}
incr :: (KnownSymbol s, Get xs s ~ Just (StringOf Integer))
     => Proxy s -> Edis xs xs (Either Reply Integer)
incr key = Edis $ Hedis.incr (symbolVal key)

incrbyfloat :: (KnownSymbol s, Get xs s ~ Just (StringOf Double))
     => Proxy s -> Double -> Edis xs xs (Either Reply Double)
incrbyfloat key n = Edis $ Hedis.incrbyfloat (symbolVal key) n
\end{minted}

@@

Consider the following example: \mintinline{text}{LLEN} returns the length of
the list associated with a key, else raises a type error.

@@

We could achieve this simply by translating the semantics we want to the
domain of Boolean, with type-level boolean functions such as
\mintinline{haskell}{(&&)},
\mintinline{haskell}{(||)}, \mintinline{haskell}{Not},
\mintinline{haskell}{(==)}, etc.\footnotemark To avoid

@@

To avoid addressing the type of value (as it may not exist at all), we defined an
 auxiliary predicate \mintinline{haskell}{IsList :: Maybe * -> Bool} to replace
 the former part.

@@

With \emph{constraint kind}, a recent addition to GHC. Type constraints now has
its own kind: \mintinline{haskell}{Constraint}. That means type constraints
are not restricted to the left side of a \mintinline{haskell}{=>} anymore,
they could appear in anywhere that accepts something of kind
\mintinline{haskell}{Constraint}, and any type that has kind
\mintinline{haskell}{Constraint} can also be used as a type constraint.\footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}

As many other list-related commands also have this "List or Nothing" semantics,
we could abstract the lengthy type constraint above and give it an alias with
type synonym.

@@

The complete implementation of \mintinline{text}{LLEN} with
\mintinline{haskell}{ListOrNX} would become:

\begin{minted}[xleftmargin=1em]{haskell}
llen :: (KnownSymbol s, ListOrNX xs s)
        => Proxy s
        -> Edis xs xs (Either Reply Integer)
llen key = Edis $ Hedis.llen (symbolVal key)
\end{minted}

\section{Assertions}

Users may need to make assertions about the status of certain key-type bindings
 in a Redis program. For example, declaring the existence of a key and the type
 of its associating value. We provide these functions, which do nothing but
 fiddling with types.

\begin{minted}[xleftmargin=1em]{haskell}
declare :: (KnownSymbol s, Member xs s ~ False)
        => Proxy s
        -> Proxy x  -- type of value
        -> Edis xs (Set xs s x) ()
declare s x = Edis $ return ()

renounce :: (KnownSymbol s, Member xs s ~ True)
        => Proxy s -> Edis xs (Del xs s) ()
renounce s = Edis $ return ()

-- empty precondition
start :: Edis ’[] ’[] ()
start = Edis $ return ()
\end{minted}

\subsection{A complete example}

The following program increases the value of "A" as an integer,
and push the result of increment to list "L", and then pops it out.

\begin{minted}[xleftmargin=1em]{haskell}
main :: IO ()
main = do
    conn <- connect defaultConnectInfo
    result <- runRedis conn $ unEdis $ start
        `bind` \_ ->  declare
                        (Proxy :: Proxy "A")
                        (Proxy :: Proxy Integer)
        `bind` \_ ->  incr     (Proxy :: Proxy "A")
        `bind` \n ->  case n of
            Left  err -> lpush (Proxy :: Proxy "L") 0
            Right n   -> lpush (Proxy :: Proxy "L") n
        `bind` \_ ->  lpop     (Proxy :: Proxy "L")
    print result
\end{minted}

The syntax is pretty heavy, like the old days when there's no
do-notation\cite{history}. But if we don't need any variable bindings between
operations, we could compose these commands with a sequencing operator
\mintinline{haskell}{(>>>)}.

\begin{minted}[xleftmargin=1em]{haskell}
(>>>) :: IMonad m => m p q a -> m q r b -> m p r b

program = start
    >>> declare
            (Proxy :: Proxy "A")
            (Proxy :: Proxy Integer)
    >>> incr    (Proxy :: Proxy "A")
    >>> lpush   (Proxy :: Proxy "L") 0
    >>> lpop    (Proxy :: Proxy "L")
\end{minted}

We could even get rid of \mintinline{haskell}{(>>>)} with
 \emph{rebindable syntax}\footnote{See \url{https://downloads.haskell.org/~ghc/latest/docs/html/users_guide/syntax-extns.html#rebindable-syntax}.},
if we are so inclined.

\section{Discussions}
\paragraph{Syntax}
\paragraph{Commands with multiple inputs or outputs}
\paragraph{Redis Transactions}
\section{Conclusion and Related Work}
\paragraph{Acknowledgements}
